---
title: "Hw7 Stat 139"
author: "Callin Switzer"
date: "October 28, 2014"
output: pdf_document
---

```{r, results = "asis", echo = F, warning=F, message = F}
options(xtable.comment = FALSE)
births <- read.csv("data/ex0724.csv")

countries <- colnames(births)[2:5]

mods <- lapply(countries, function(x) summary(lm(births[[x]] ~ births$Year)))

names(mods) <- countries


require(xtable)
x <- "Denmark"

foo <- lm(births[[x]]~births$Year)
sigmaHat <- sqrt(sum((foo$residuals)^2)/length(foo$residuals))
#summary(foo)
#print(xtable(foo))
```

# #2 (#26a)
Below are the tables for the linear regression for each country.  I've also calculatedd the $\hat{\sigma}$ (sigmaHat) for each country and printed it under its respective table.  The values are all the same as Diaplay 7.17 in the book


```{r, results = "asis", echo = F}
options("scipen"=10, "digits"=4)
for(i in 1:4){
     sigmaHat <-  sqrt(sum((mods[[i]]$residuals)^2)/(length(mods[[i]]$residuals) - 2))
     sigmaHat <- round(sigmaHat, digits = 5)
     capt = (paste("sigmaHat =", sigmaHat, "for", countries[i]))
     print(xtable(mods[[i]], digits = c(10, 10, 10, 10, 10),  
                  caption = capt), floating = F)
     print(paste(capt))
}
```

# 26b
The t-statistics are shown in the tables above. There is evidence that the proportion of male birth is truly declining.  The p-values are highly significant for Netherlands, Candada, and USA, and somewhat significant for Denmark.


# 26c
The t-statistic is the ratio of the coefficient estimates to their standard errors.  The US can have an extreme t-statistic because the standard error is very small, relative to the estimate.  This makes the t-statistic more extreme.




# 26d
The standard error for the estimated slope is smaller in the US than Canada because the points fall closer to the line. Standard error of $\beta_1$ depends on $\hat{\sigma}$, which becomes larger when the sum of squared residuals gets larger. The SE($\beta_1$) also depends on the sample variance of the x-values, but those should be the same, since the US and Canada were sampled during the same years. 



# Code
```{r, eval = F}
options(xtable.comment = FALSE)
births <- read.csv("data/ex0724.csv")

countries <- colnames(births)[2:5]

mods <- lapply(countries, function(x) summary(lm(births[[x]] ~ births$Year)))

names(mods) <- countries


require(xtable)
x <- "Denmark"

foo <- lm(births[[x]]~births$Year)
sigmaHat <- sqrt(sum((foo$residuals)^2)/length(foo$residuals))
#summary(foo)
#print(xtable(foo))

options("scipen"=10, "digits"=4)
for(i in 1:4){
     sigmaHat <-  sqrt(sum((mods[[i]]$residuals)^2)/(length(mods[[i]]$residuals) - 2))
     sigmaHat <- round(sigmaHat, digits = 5)
     capt = (paste("sigmaHat =", sigmaHat, "for", countries[i]))
     print(xtable(mods[[i]], digits = c(10, 10, 10, 10, 10),  
                  caption = capt), floating = F)
     print(paste(capt))
}
```